Paper Analysis

### 2025-01-13 Paper_WriteUp/Paper_Plots.qmd
add Adam's code for figure 2

```{r figure 2}

## to be added

>>>>

```{r figure 2}
## Adam's code for this figure

gm_players_elo <- list()
gm_session_lengths <- list()
gm_time_diffs <- list()
gm_time_diffs_10 <- list()

gm_players <- unique(bullet_data_gm$Username)

for (i in 1:length(gm_players)) {
  tidy_games_gms <- bullet_data_gm |>
    filter(Username == gm_players[i]) |> # select only one player
    mutate(UTCDateTime = ymd_hms(paste(UTCDate, UTCTime))) |> # convert date-time to lubridate type
    arrange(UTCDateTime) |> # order date-time
    mutate(time_diff = UTCDateTime - lag(UTCDateTime)) # time difference between games

  ## A check to remove duplicated games, if they occur
  # cat("Dupes:", sum(duplicated(tidy_games_gms[1:9])), "\n", sep = " ")
  # # some players have all their games double recorded for some reason
  # tidy_games_gms <- tidy_games_gms[!duplicated(tidy_games_gms[1:9]), ] # remove duplicate rows

  # get players average elo across all games
  gm_players_elo[[gm_players[i]]] <- mean(c(
    tidy_games_gms$WhiteElo[gm_players[i] == tidy_games_gms$White],
    tidy_games_gms$BlackElo[gm_players[i] == tidy_games_gms$Black]
  ))

  # store time diffs between games
  gm_time_diffs[[gm_players[i]]] <- difftime(tidy_games_gms$UTCDateTime,
    lag(tidy_games_gms$UTCDateTime),
    units = "secs"
  ) # time between current game and previous
  gm_time_diffs_10[[gm_players[i]]] <- difftime(tidy_games_gms$UTCDateTime,
    lag(tidy_games_gms$UTCDateTime, 10),
    units = "secs"
  ) # time between current game and 10th game ago
}
# store in df for plotting
gm_time_diff_df <- data.frame("time_diff" = unlist(gm_time_diffs))
gm_time_diff_df_10 <- data.frame("time_diff" = unlist(gm_time_diffs_10))

## Same distributions as amateurs, so again proceed by assuming all 
## games played within 300 seconds (5 min) are in the same session
for (i in 1:length(gm_players)) {
  tidy_games_gms <- bullet_data_gm |>
    filter(Username == gm_players[i]) |> # select only one player
    mutate(UTCDateTime = ymd_hms(paste(UTCDate, UTCTime))) |> # convert date-time to lubridate type
    arrange(UTCDateTime) |> # order date-time
    mutate(time_diff = UTCDateTime - lag(UTCDateTime)) # time difference between games

  length_of_sessions <- which(tidy_games_gms$time_diff > 300) - 
    lag(which(tidy_games_gms$time_diff > 300), default = 0)
  # game id of first game of session minus game id of first game of previous session

  gm_session_lengths[[gm_players[i]]] <- length_of_sessions # store
}
# store in df for plotting
gm_session_lengths_df <- data.frame(session_length = unlist(gm_session_lengths))

log_time_diff_plot_gm <- ggplot(gm_time_diff_df, aes(x = as.numeric(time_diff))) +
  geom_histogram() +
  scale_x_continuous(
    trans = scales::log_trans(),
    breaks = c(1, 60, 300, 3600, 86400, 31536000),
    labels = c("1 sec", "1 min", "5 min", "1 hour", "1 day", "1 year")
  ) +
  xlab("Time Between Games (Log Scale)") +
  geom_vline(xintercept = 300, colour = "red") +
  ylab("Frequency") #+ ggtitle("Time Difference Between Current Game and Previous Game for Amateurs")
fig2a <- log_time_diff_plot_gm + theme_single()

ggsave(plot = fig2a, 
       filename = paste0(here("Paper_WriteUp", "paper_figures",
                              "time_between_games_gm.png")), dpi = 600,
       height = 5, width = 7)

session_length_plot_gm <- ggplot(gm_session_lengths_df, aes(x = session_length)) +
  geom_histogram() +
  scale_x_continuous(
    trans = pseudo_log_trans(sigma = 2.36),
    breaks = c(1, 5, 10, 100),
    labels = c("1 game", "5 games", "10 games", "100 games"),
    limits = c(0, 389)
  ) +
  xlab("Number of Games in a Session (Log Scale)") +
  ylab("Frequency") #+ ggtitle("Number of Games per Session for GMs")
fig2b <- session_length_plot_gm + theme_single()

ggsave(plot = fig2b, 
       filename = paste0(here("Paper_WriteUp", "paper_figures",
                              "session_length_gm.png")), dpi = 600,
       height = 5, width = 7)
```
